---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(longears)
```
# longears

This repository contains an early stage proof-of-concept RabbitMQ client for R.
It wraps the [reference C library](https://github.com/alanxz/rabbitmq-c) instead
of taking the rJava approach of the only other [existing (but abandoned)](https://r-forge.r-project.org/projects/r-message-queue/)
package.

## Installation

You will need `librabbitmq`, otherwise the installation will fail. On Ubuntu, run

```shell
$ apt install librabbitmq-dev
```

The package is only available from GitHub for now, so you can install it with

``` r
# install.packages("devtools")
devtools::install_github("atheriel/longears")
```

## Usage

You will need to have a local RabbitMQ server running with the default
settings to test this.

```shell
$ # apt install rabbitmq-server
$ systemctl start rabbitmq-server
$ rabbitmqctl status
```

To connect to the server (with default settings):

```{r basic-connect}
conn <- amqp_connect()
conn
```

To send messages to a queue:

```{r basic-publish}
amqp_declare_queue(conn, "my_queue")
amqp_publish(conn, "my_queue", "message")
amqp_publish(conn, "my_queue", "second message")
```

Check if your messages are going into the queue:

```shell
$ rabbitmqctl list_queues
```

And to pull messages back into R:

```{r basic-get}
amqp_get(conn, "my_queue")
amqp_get(conn, "my_queue")
amqp_get(conn, "my_queue")
```

Afterwards you can delete the queue and disconnect from the server:

```{r basic-disconnect}
amqp_delete_queue(conn, "my_queue")
amqp_disconnect(conn)
conn
```

And check that the connection is closed:

```shell
$ rabbitmqctl list_connections
```

## License

The package is licensed under the GPL, version 2 or later.
