---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(longears)
```
# longears

<!-- badges: start -->
[![travis-ci build status](https://travis-ci.org/atheriel/longears.svg?branch=master)](https://travis-ci.org/atheriel/longears)
<!-- badges: end -->

This repository contains an early stage RabbitMQ client for R. It wraps the
[reference C library](https://github.com/alanxz/rabbitmq-c).

[RabbitMQ](https://www.rabbitmq.com/) is a popular and performant open-source
message broker used to build highly distributed and asynchronous network
topologies. This package may be of interest to you if you wish to have R speak
to your organization's existing RabbitMQ servers; if you simply need a message
queue library, you may be better off with [**txtq**](https://github.com/wlandau/txtq),
[**litq**](https://github.com/r-lib/liteq), or the ZeroMQ package [**rzmq**](https://github.com/ropensci/rzmq).

The package implements a large portion of the Advanced Message Queuing Protocol
(AMQP) used by RabbitMQ , and the API largely reflects [the protocol itself](https://www.rabbitmq.com/amqp-0-9-1-reference.html).
However, `longears` is not a true low-level package: it abstracts away many
details of AMQP for end users. See [Limitations](#Limitations) for details.

## Installation

You will need `librabbitmq`, otherwise the installation will fail. On Ubuntu, run

```shell
$ apt install librabbitmq-dev
```

The package is only available from GitHub for now, so you can install it with

``` r
# install.packages("devtools")
devtools::install_github("atheriel/longears")
```

## Basic Usage

If you are not already familiar with RabbitMQ and the message/queue/binding/exchange
terminology, I suggest their excellent [conceptual overview](https://www.rabbitmq.com/tutorials/amqp-concepts.html).

You will also need to have a local RabbitMQ server running with the default
settings to follow this guide.

```shell
$ # apt install rabbitmq-server
$ systemctl start rabbitmq-server
$ rabbitmqctl status
```

First, connect to the server (with the default settings):

```{r basic-connect}
conn <- amqp_connect()
conn
```

Create an exchange to route messages and a couple of queues to store them:

```{r basic-exchange}
amqp_declare_exchange(conn, "my.exchange", type = "fanout")
amqp_declare_queue(conn, "my.queue1")
amqp_declare_queue(conn, "my.queue2")
amqp_bind_queue(conn, "my.queue1", "my.exchange", routing_key = "#")
amqp_bind_queue(conn, "my.queue2", "my.exchange", routing_key = "#")
```

You can also set up a consumer for one of these queues with a callback:

```{r basic-consume}
received <- 0
consumer <- amqp_consume(conn, "my.queue2", function(msg) {
  received <<- received + 1
})
```

Now, send a few messages to this exchange:

```{r basic-publish}
amqp_publish(conn, "first", exchange = "my.exchange", routing_key = "#")
amqp_publish(conn, "second", exchange = "my.exchange", routing_key = "#")
```

Check if your messages are going into the queues:

```shell
$ rabbitmqctl list_queues
```

You can use `amqp_get()` to pull individual messages back into R:

```{r basic-get}
amqp_get(conn, "my.queue1")
amqp_get(conn, "my.queue1")
amqp_get(conn, "my.queue1")
```

Or you can use `amqp_listen()` to run consumer callbacks:

```{r basic-listen}
amqp_listen(conn, timeout = 1)
received
```

To clean things up, delete the queues, the exchange, and disconnect from the
server.

```{r basic-disconnect}
amqp_delete_queue(conn, "my.queue1")
amqp_delete_queue(conn, "my.queue2")
amqp_delete_exchange(conn, "my.exchange")
amqp_disconnect(conn)
conn
```

And check that the connection is closed:

```shell
$ rabbitmqctl list_connections
```

## Limitations

Some AMQP features are not present, notably transaction support (which there
are no plans to implement). Others are not exposed through the API but are
handled internally according to best practices: channels, message
acknowledgements, and prefetch counts, among others. The current design goal of
the package is to keep these details from the eyes of users, insofar as that is
possible.

If you have need of an AMQP feature (or RabbitMQ extension) that is not
currently available or accessible, please consider filing an issue explaining
the use case you have in mind.

## License

The package is licensed under the GPL, version 2 or later.
