---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(longears)
```
# longears

This repository contains an early stage RabbitMQ client for R. It wraps the
[reference C library](https://github.com/alanxz/rabbitmq-c).

[RabbitMQ](https://www.rabbitmq.com/) is a popular and performant open-source
message broker used to build highly distributed and asynchronous network
topologies. This package may be of interest to you if you wish to have R speak
to your organization's existing RabbitMQ servers; if you simply need a message
queue library, you may be better off with [**txtq**](https://github.com/wlandau/txtq),
[**litq**](https://github.com/r-lib/liteq), or the ZeroMQ package [**rzmq**](https://github.com/ropensci/rzmq).

The package implements a growing subset of the Advanced Message Queuing Protocol
(AMQP) used by RabbitMQ (see [Limitations](#Limitations) for details), and the
API largely reflects [the protocol itself](https://www.rabbitmq.com/amqp-0-9-1-reference.html).

## Installation

You will need `librabbitmq`, otherwise the installation will fail. On Ubuntu, run

```shell
$ apt install librabbitmq-dev
```

The package is only available from GitHub for now, so you can install it with

``` r
# install.packages("devtools")
devtools::install_github("atheriel/longears")
```

## Usage

You will need to have a local RabbitMQ server running with the default
settings to test this.

```shell
$ # apt install rabbitmq-server
$ systemctl start rabbitmq-server
$ rabbitmqctl status
```

First, connect to the server (with default settings):

```{r basic-connect}
conn <- amqp_connect()
conn
```

Create an exchange to route messages and a queue to store them:

```{r basic-exchange}
amqp_declare_exchange(conn, "my.exchange")
amqp_declare_queue(conn, "my.queue")
amqp_bind_queue(conn, "my.queue", "my.exchange", routing_key = "#")
```

Now, send a message to this exchange:

```{r basic-publish}
amqp_publish(conn, "#", "message", exchange = "my.exchange")
amqp_publish(conn, "#", "second message", exchange = "my.exchange")
```

Check if your messages are going into the queue:

```shell
$ rabbitmqctl list_queues
```

And to pull messages back into R:

```{r basic-get}
amqp_get(conn, "my.queue")
amqp_get(conn, "my.queue")
amqp_get(conn, "my.queue")
```

Afterwards you can delete the queue and disconnect from the server:

```{r basic-disconnect}
amqp_delete_queue(conn, "my.queue")
amqp_delete_exchange(conn, "my.exchange")
amqp_disconnect(conn)
conn
```

And check that the connection is closed:

```shell
$ rabbitmqctl list_connections
```

## Limitations

There are many AMQP features missing at present but that will be added in the
future. An incomplete list is as follows:

- Additional Basic methods.
- Additional Basic properties.
- Message headers.
- Support for "table" arguments to e.g. exchange declarations.

In addition, the API does not at present implement any asynchronous
functionality. Most notably, there is no `consume` support, although it is my
hope a design for one can be worked out in the future.

## License

The package is licensed under the GPL, version 2 or later.
